Arrows for Parallel Computations

### 2018-01-05 paper/content/benchmarks.tex
martin remarks

In the following section, we describe the benchmarks of our parallel DSL and algorithmic skeletons conducted.
We start by explaining the hardware and software stack and also elaborate on which benchmarks programs were used
and also which parallel Haskell were used in which setting and why. Before we go into detail on the benchmarks we
also address the issue of poor performance gains when using hyper-threading and our reasoning behind not including
hyper-threaded cores in our benchmarks. Finally, we show that PArrows hold up in terms of performance when compared to the original parallel Haskells used as backends in this paper, starting with the shared-memory variants (GpH, |Par| Monad and Eden CP) and concluding with Eden as a distributed backend.

@@

The \ensuremath{\Conid{Par}} monad and GpH -- in its multicore version \cite{Marlow2009} --  can be executed on shared memory machines only.
Although GpH is available on distributed memory
clusters, and newer distributed memory Haskells such as HdpH exist,
current support of distributed memory in PArrows is limited to
Eden. We used the MPI backend of Eden in a distributed memory
setting. However, for shared memory Eden features a ``CP'' backend
that merely copies the memory blocks between distributed heaps. In
this mode, Eden still operates in the ``nothing shared'' setting, but
is adapted better to multicore machines. We label this version of Eden
in the plots as ``Eden~CP''.

@@

In preliminary tests, the PArrows version of \rmtest on a single node of the Glasgow cluster
showed almost linear speedup on up to 16 shared-memory cores (as supplementary materials show). The speedup
of 64-task PArrows/Eden at 16 real cores version was 13.65 giving a parallel
efficiency of 85.3\%. However, if we increased the number of
requested cores to 32 -- \ie if we use hyper-threading on 16 real
cores -- the speedup did not increase that well. It was merely 15.99
for 32 tasks with PArrows/Eden. This was worse for other backends.  As
for 64 tasks, we obtained a speedup of 16.12 with PArrows/Eden at 32
hyper-threaded cores and only 13.55 with PArrows/GpH. The parallel efficiencies were 50.4\% and 42.3\%, respectively. The Eden
version used here was Eden~CP, the \enquote{share nothing} SMP build.

In the distributed memory setting the same effect ensues. We obtain
plummeting speedup of 124.31 at 512 hyper-threaded cores, whereas it was
213.172 for 256 real cores. Apparently, hyper-threading in the Glasgow
cluster fails to execute two parallel Haskell processes with full-fledged
parallelism.

\mbcomment{fix this when benchmarks are redone}
With this in mind, the performance measurements in the Section~\ref{sec:benchmarkResults} use only real cores as the purpose of this paper is to show the performance of PArrows and not to investigate parallel performance with hyper-threading in more detail.

@@

We compared the overhead of our PArrows implementations to the original parallel Haskells. We basically implemented the test programs both in PArrows and in traditional style, benchmarked both and compare the mean overhead, i.e. the \textit{relative} difference between both benchmarks with same settings. The error margin of the time measurements, supplied by criterion package, is computed together with the mean overhead computation.

Quite often the zero value lies in the error margin of the mean overhead. This means that though we have measured some difference (against or even in favour of PArrows), it could be merely the error margin of the benchmarks and the difference might not be existent. We are mostly interested in the cases where above issue does not persist, we call them \emph{significant}. We often denote the error margin with $\pm$ after the mean overhead value.

@@

\paragraph{Speedup}
The \rmtest benchmark showed almost linear speedup for both 32 and 64 tasks, the performance is slightly better in the latter case: 13.7 at 16 cores for input $2^{11213}-1$ and 64 tasks in the best case scenario with Eden CP. The performance of the \sudokutest benchmark merely reaches a speedup of 9.19 (GpH), 8.78 (Par Monad), 8.14 (Eden CP) for 16 cores and 1000 Sudokus. In contrast to Rabin–Miller, here the |GpH| Monad backend seems to be the best of all, while Rabin–Miller profited most from Eden CP (i.e., Eden with direct memory copy) implementation of PArrows. Gentleman on shared memory has a plummeting speedup curve with GpH and |Par| Monad and logarithmically increasing speedup for Eden-based version. The latter reached a speedup of 6.56 at 16 cores.

@@

For Sudoku the situation is slightly different. There is a minimal significant ($-1.4\% \; \pm 1.2\%$ at 8 cores) speed improvement with PArrows Eden~CP version when compared with the base Eden CP benchmark. However, with increasing number of cores the error margin reaches zero again: $-1.6\% \; \pm 5.0\%$ at 16 cores. The |Par| Monad shows a similar development, \eg with $-1.95\% \; \pm 0.64\%$ at 8 cores. The GpH version shows both a significant speed \textit{improvement} from PArrows of $-4.2\% \; \pm 0.26\%$ and a minor overhead of $0.87\% \; \pm 0.70\%$.

>>>>

In the following section, we describe the benchmarks we conducted of our parallel DSL and algorithmic skeletons.
We start by explaining the hardware and software stack and also elaborate on which benchmarks programs and parallel Haskells were used in which setting and why. Before we go into detail on the benchmarks we
also shortly address hyper-threading and why we do not use it in our benchmarks. Finally, we show that PArrows hold up in terms of performance when compared to the original parallel Haskells used as backends in this paper, starting with the shared-memory variants (GpH, |Par| Monad and Eden CP) and concluding with Eden as a distributed backend.

@@

The \ensuremath{\Conid{Par}} monad and GpH -- in its multicore version \cite{Marlow2009} --  can be executed on shared memory machines only.
Although GpH is available on distributed memory
clusters, and newer distributed memory Haskells such as HdpH exist,
current support of distributed memory in PArrows is limited to
Eden. We used the MPI backend of Eden in a distributed memory
setting. However, for shared memory Eden features a \enquote{CP} backend
that merely copies the memory blocks between distributed heaps. In
this mode, Eden still operates in the \enquote{nothing shared} setting, but
is adapted better to multicore machines. We call this version of Eden
\enquote{Eden~CP}.

@@

In preliminary tests, the PArrows version of \rmtest on a single node of the Glasgow cluster
showed almost linear speedup on up to 16 shared-memory cores (as supplementary materials show). The speedup
of 64-task PArrows/Eden at 16 real cores version was 13.65 giving a parallel
efficiency of 85.3\%. However, if we increased the number of
requested cores to 32 -- \ie if we use hyper-threading on 16 real
cores -- the speedup did not increase that well. It was merely 15.99
for 32 tasks with PArrows/Eden. This was worse for other backends.  As
for 64 tasks, we obtained a speedup of 16.12 with PArrows/Eden at 32
hyper-threaded cores and only 13.55 with PArrows/GpH.

While this shows that hyper-threading can be of benefit in scenarios similar to the ones presented in the benchmarks, we only use real cores for the performance measurements in Section~\ref{sec:benchmarkResults} as the purpose of this paper is to show the performance of PArrows and not to investigate parallel behaviour with hyper-threading.

@@

We compared the overhead of our PArrows implementations to the original parallel Haskells. We basically implemented the test programs both in PArrows and in traditional style, benchmarked both and will now compare the mean overhead, i.e. the \textit{relative} difference between both benchmarks with same settings. The error margin of the time measurements, supplied by criterion package, is computed together with the mean overhead computation.

Quite often the zero value lies in the error margin of the mean overhead. This means that even though we have measured some difference (against or even in favour of PArrows), it could be merely the error margin of the benchmarks and the difference might not be existent. We are mostly interested in the cases where above issue does not persist, we call them \emph{significant}. We often denote the error margin with $\pm$ after the mean overhead value.

@@

\paragraph{Speedup}
The \rmtest benchmark showed almost linear speedup for both 32 and 64 tasks, the performance is slightly better in the latter case: 13.7 at 16 cores for input $2^{11213}-1$ and 64 tasks in the best case scenario with Eden CP. The performance of the \sudokutest benchmark merely reaches a speedup of 9.19 (GpH), 8.78 (Par Monad), 8.14 (Eden CP) for 16 cores and 1000 Sudokus. In contrast to Rabin–Miller, here the |GpH| backend seems to be the best of all, while Rabin–Miller profited most from Eden CP (i.e., Eden with direct memory copy) implementation of PArrows. Gentleman on shared memory has a plummeting speedup curve with GpH and |Par| Monad and logarithmically increasing speedup for the Eden-based version. The latter reached a speedup of 6.56 at 16 cores.

@@

For Sudoku the situation is slightly different. There is a minimal significant ($-1.4\% \; \pm 1.2\%$ at 8 cores) speed improvement with PArrows Eden~CP version when compared with the base Eden CP benchmark. However, with increasing number of cores the error margin reaches zero again: $-1.6\% \; \pm 5.0\%$ at 16 cores. The |Par| Monad shows a similar development, \eg with $-1.95\% \; \pm 0.64\%$ at 8 cores. The GpH version shows both a significant speed \textit{improvement} from PArrows of $-4.2\% \; \pm 0.26\%$ and a minor overhead of $0.87\% \; \pm 0.70\%$. \mbcomment{Widerspricht sich das nicht?}

### 2018-01-05 paper/content/topologySkeletons.tex
martin remarks

%\FloatBarrier
\subsection{Topological Skeletons}
\label{sec:topology-skeletons}
Even though many algorithms can be expressed by parallel maps, some problems require more sophisticated skeletons. The Eden library leverages this problem and already comes with more predefined skeletons\footnote{Available on Hackage: \url{https://hackage.haskell.org/package/edenskel-2.1.0.0/docs/Control-Parallel-Eden-Topology.html}.}, among them a |pipe|, a |ring|, and a |torus| implementations \citep{Eden:SkeletonBookChapter02}. These seem like reasonable candidates to be ported to our Arrow-based parallel Haskell. We aim to showcase that we can express more sophisticated skeletons with Parallel Arrows as well.

If we used the original definition of |parEvalN|, however, these skeletons would produce an infinite loop with the GpH and Par Monad backends which during runtime would result in the program crashing. This materialises with the usage of |loop| of the |ArrowLoop| type class and we think that this is due to difference of how evaluation is done in these backends when compared to Eden. An investigation of where this difference comes from is beyond the scope of this work, we only provide a workaround for these types of skeletons as such skeletons probably are not of much importance outside of a distributed memory environment. However our workaround enables users of the DSL to test their code within a shared memory setting even though other skeletons would be better suited to be run with them.
The idea of the fix is to provide a |ArrowLoopParallel| typeclass that has two functions - |loopParEvalN| and |postLoopParEvalN| where the first is to be used inside an |loop| construct while the latter will be used right outside of the |loop|. This way we can delegate to the actual |parEvalN| in the spot where the backend supports it.
\begin{code}
class ArrowParallel arr a b conf => ArrowLoopParallel arr a b conf where
    loopParEvalN :: conf -> [arr a b] -> arr [a] [b]
    postLoopParEvalN :: conf -> [arr a b] -> arr [a] [b]
\end{code}
As Eden has no problems with the looping skeletons, we use this instance:
\begin{code}
instance (ArrowChoice arr, ArrowParallel arr a b Conf) => ArrowLoopParallel arr a b Conf where
    loopParEvalN = parEvalN
    postLoopParEvalN _ = evalN
\end{code}
As the backends for the |Par| Monad and GpH have problems with |parEvalN| inside of |loop| their respective instances for |ArrowLoopParallel| look like this:
\begin{code}
instance (ArrowChoice arr, ArrowParallel arr a b (Conf b)) => ArrowLoopParallel arr a b (Conf b) where
    loopParEvalN _ = evalN
    postLoopParEvalN = parEvalN
\end{code}

@@

The parallel |pipe| skeleton is semantically equivalent to folding over a list |[arr a a]| of arrows with |>>>|, but does this in parallel, meaning that the arrows do not have to reside on the same thread/machine. We implement this skeleton using the |ArrowLoop| type class which gives us the |loop :: arr (a, b) (c, b) -> arr a c| combinator which allows us to express recursive fix-point computations in which output values are fed back as input. For example %this
\begin{code}
loop (arr (\(a, b) -> (b, a:b)))
\end{code}
which is the same as
\begin{code}
loop (arr snd &&& arr (uncurry (:)))
\end{code}
defines an arrow that takes its input |a| and converts it into an infinite stream |[a]| of it. Using this to our advantage gives us a first draft of a pipe implementation (Fig.~\ref{fig:pipeSimple}) by plugging in the parallel evaluation call |evalN conf fs| inside the second argument of |&&&| and then only picking the first element of the resulting list with |arr last|.

\begin{figure}[t]
\begin{code}
pipeSimple :: (ArrowLoop arr, ArrowLoopParallel arr a a conf) =>
	conf -> [arr a a] -> arr a a
pipeSimple conf fs =
    loop (arr snd &&&
        (arr (uncurry (:) >>> lazy) >>> loopParEvalN conf fs)) >>>
    arr last
\end{code}
\caption{A~first implementation of the |pipe| skeleton expressed with Parallel Arrows. Note that the use of |lazy| (Fig.~\ref{fig:edenlazyrightrotate}) is essential as without it programs using this definition would never halt. We need to enforce that the evaluation of the input |[a]| terminates before passing it into |evalN|.}
\label{fig:pipeSimple}

@@

\begin{code}
pipe :: (ArrowLoop arr, ArrowLoopParallel arr (fut a) (fut a) conf,
	Future fut a conf) =>
	conf -> [arr a a] -> arr a a
pipe conf fs = unliftFut conf (pipeSimple conf (map (liftFut conf) fs))

@@

Note that extensive use of |pipe2| over |pipe| with a hand-written combination data type will probably result in worse performance because of more communication overhead from the many calls to |parEvalN| inside of |evalN|. Nonetheless, we can define a version of parallel piping operator |parcomp|, which is semantically equivalent to |>>>| similarly to other parallel syntactic sugar from Appendix~\ref{syntacticSugar}.

@@

We can rewrite this functionality easily with the use of |loop| as the definition of the node function, |arr (i, r) (o, r)|, after being transformed into an arrow, already fits quite neatly into the |loop|'s |arr (a, b) (c, b) -> arr a c|. In each iteration we start by rotating the intermediary input from the nodes |[fut r]| with |second (rightRotate >>> lazy)| (Fig.~\ref{fig:edenlazyrightrotate}). Similarly to the |pipe| from Section~\ref{sec:pipe} (Fig.~\ref{fig:pipeSimple}), we have to feed the intermediary input into our |lazy| (Fig.~\ref{fig:edenlazyrightrotate}) arrow here, or the evaluation would fail to terminate. The reasoning is explained by \citet{Loogen2012} as a demand problem.

Next, we zip the resulting |([i], [fut r])| to |[(i, fut r)]| with |arr (uncurry zip)| so we can feed that into our input arrow |arr (i, r) (o, r)|, which we transform into |arr (i, fut r) (o, fut r)| with |second get >>> f >>> second put| before converting it into a parallel arrow |arr [(i, fut r)] [(o, fut r)]| with the help of |repeat| and |loopParEvalN| to get an output list |[(o, fut r)]|. Finally we unzip this list into |([o], [fut r])|. Plugging this arrow |arr ([i], [fut r]) ([o], fut r)| into the definition of |loop| from earlier gives us |arr [i] [o]|, our ring arrow (Fig.~\ref{fig:ringFinal}). To make sure this algorithm has speedup on shared-memory machines as well, we pass the result of this arrow to |postLoopParEvalN conf (repeat (arr id))|.
This combinator can, for example, be used to calculate the shortest paths in a graph using Warshall's algorithm.
%Further details on this can be found in \cite{eden_cefp}.

>>>>

%\FloatBarrier
\subsection{Topological Skeletons}
\label{sec:topology-skeletons}
Even though many algorithms can be expressed by parallel maps, some problems require more sophisticated skeletons. The Eden library leverages this problem and already comes with more predefined skeletons\footnote{Available on Hackage: \url{https://hackage.haskell.org/package/edenskel-2.1.0.0/docs/Control-Parallel-Eden-Topology.html}.}, among them a |pipe|, a |ring|, and a |torus| implementations \citep{Eden:SkeletonBookChapter02}. These seem like reasonable candidates to be ported to our Arrow-based parallel Haskell. We aim to showcase that we can express more sophisticated skeletons with parallel Arrows as well.

If we used the original definition of |parEvalN|, however, these skeletons would produce an infinite loop with the GpH and Par Monad backends which during runtime would result in the program crashing. This materialises with the usage of |loop| of the |ArrowLoop| type class and we think that this is due to difference of how evaluation is done in these backends when compared to Eden. An investigation of why this difference exists is beyond the scope of this work, we only provide a workaround for these types of skeletons as such they probably are not of much importance outside of a distributed memory environment. However our workaround enables users of the DSL to test their code within a shared memory setting even though other skeletons would be better suited to be run with them.
The idea of the fix is to provide a |ArrowLoopParallel| typeclass that has two functions - |loopParEvalN| and |postLoopParEvalN| where the first is to be used inside an |loop| construct while the latter will be used right outside of the |loop|. This way we can delegate to the actual |parEvalN| in the spot where the backend supports it.
\begin{code}
class ArrowParallel arr a b conf =>
	ArrowLoopParallel arr a b conf where
    loopParEvalN :: conf -> [arr a b] -> arr [a] [b]
    postLoopParEvalN :: conf -> [arr a b] -> arr [a] [b]
\end{code}
As Eden has no problems with the looping skeletons, we use this instance:
\begin{code}
instance (ArrowChoice arr, ArrowParallel arr a b Conf) =>
	ArrowLoopParallel arr a b Conf where
    loopParEvalN = parEvalN
    postLoopParEvalN _ = evalN
\end{code}
As the backends for the |Par| Monad and GpH have problems with |parEvalN| inside of |loop| their respective instances for |ArrowLoopParallel| look like this:
\begin{code}
instance (ArrowChoice arr, ArrowParallel arr a b (Conf b)) =>
	ArrowLoopParallel arr a b (Conf b) where
    loopParEvalN _ = evalN
    postLoopParEvalN = parEvalN
\end{code}

@@

The parallel |pipe| skeleton is semantically equivalent to folding over a list |[arr a a]| of arrows with |>>>|, but does this in parallel, meaning that the arrows do not have to reside on the same thread/machine. We implement this skeleton using the |ArrowLoop| type class which gives us the |loop :: arr (a, b) (c, b) -> arr a c| combinator which allows us to express recursive fix-point computations in which output values are fed back as input. For example %this
\begin{code}
loop (arr (\(a, b) -> (b, a:b)))
\end{code}
which is the same as
\begin{code}
loop (arr snd &&& arr (uncurry (:)))
\end{code}
defines an arrow that takes its input |a| and converts it into an infinite stream |[a]| of it. Using |loop| to our advantage gives us a first draft of a pipe implementation (Fig.~\ref{fig:pipeSimple}) by plugging in the parallel evaluation call |evalN conf fs| inside the second argument of |&&&| and then only picking the first element of the resulting list with |arr last|.

\begin{figure}[t]
\begin{code}
pipeSimple :: (ArrowLoop arr, ArrowLoopParallel arr a a conf) =>
	conf -> [arr a a] -> arr a a
pipeSimple conf fs =
    loop (arr snd &&&
        (arr (uncurry (:) >>> lazy) >>> loopParEvalN conf fs)) >>>
    arr last
\end{code}
\caption{A~first implementation of the |pipe| skeleton expressed with parallel Arrows. Note that the use of |lazy| (Fig.~\ref{fig:edenlazyrightrotate}) is essential as without it programs using this definition would never halt. We need to enforce that the evaluation of the input |[a]| terminates before passing it into |evalN|.}
\label{fig:pipeSimple}
\end{figure}

@@

\begin{figure}[t]
\begin{code}
pipe :: (ArrowLoop arr, ArrowLoopParallel arr (fut a) (fut a) conf,
	Future fut a conf) =>
	conf -> [arr a a] -> arr a a
pipe conf fs = unliftFut conf (pipeSimple conf (map (liftFut conf) fs))

@@

Note that extensive use of |pipe2| over |pipe| with a hand-written combination data type will probably result in worse performance because of more communication overhead from the many calls to |parEvalN| inside of |evalN|. Nonetheless, we can define a parallel piping operator |parcomp|, which is semantically equivalent to |>>>| similarly to other parallel syntactic sugar from Appendix~\ref{syntacticSugar}.

@@

We can rewrite this functionality easily with the use of |loop| as the definition of the node function, |arr (i, r) (o, r)|, after being transformed into an arrow, already fits quite neatly into |loop|'s signature: |arr (a, b) (c, b) -> arr a c|. In each iteration we start by rotating the intermediary input from the nodes |[fut r]| with |second (rightRotate >>> lazy)| (Fig.~\ref{fig:edenlazyrightrotate}). Similarly to the |pipe| from Section~\ref{sec:pipe} (Fig.~\ref{fig:pipeSimple}), we have to feed the intermediary input into our |lazy| (Fig.~\ref{fig:edenlazyrightrotate}) arrow here, or the evaluation would fail to terminate. The reasoning is explained by \citet{Loogen2012} as a demand problem.

Next, we zip the resulting |([i], [fut r])| to |[(i, fut r)]| with |arr (uncurry zip)|. We then feed this into our parallel Arrow |arr [(i, fut r)] [(o, fut r)]| obtained by transforming our ipnut arrow |f :: arr (i, r) (o, r)| into |arr (i, fut r) (o, fut r)| before |repeat|ing and lifting it with |loopParEvalN|. Finally we unzip the output list |[(o, fut r)]| list into |([o], [fut r])|.

Plugging this arrow |arr ([i], [fut r]) ([o], fut r)| into the definition of |loop| from earlier gives us |arr [i] [o]|, our ring arrow (Fig.~\ref{fig:ringFinal}). To make sure this algorithm has speedup on shared-memory machines as well, we pass the result of this arrow to |postLoopParEvalN conf (repeat (arr id))|.
This combinator can, for example, be used to calculate the shortest paths in a graph using Warshall's algorithm.
%Further details on this can be found in \cite{eden_cefp}.

### 2018-01-05 paper/content/utilityFunctions.tex
martin remarks

Finally, with the help of |mapArr| (Fig.~\ref{fig:mapArr}), we can define |zipWithArr|  (Fig.~\ref{fig:zipWithArr}) that lifts any Arrow |arr (a, b) c| to an Arrow |arr ([a], [b]) [c]|.
\begin{figure}[h]
\begin{code}
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr $ \(as, bs) -> zipWith (,) as bs) >>> mapArr f
\end{code}
\caption{|zipWith| over Arrows.} %$
\label{fig:zipWithArr}
\end{figure}
 %$ %% formatting

@@

\section{Syntactic Sugar} \label{syntacticSugar}
Finally, we also give the definitions for some syntactic sugar for PArrows, namely |***| and |&&&|.
For basic arrows, we have the |***| combinator (Fig.~\ref{fig:syntacticSugarArrows}) which allows us to combine two arrows |arr a b| and |arr c d| into an arrow |arr (a, c) (b, d)| which does both computations at once. This can easily be translated into a parallel version |***| with the use of |parEval2|, but for this we require a backend which has an implementation that does not require any configuration (hence the |()| as the |conf| parameter% in Fig.~\ref{fig:par***}
):
% \begin{figure}[h]
\begin{code}
(|***|) :: (ArrowChoice arr, ArrowParallel arr (Either a c) (Either b d) ())) =>
	arr a b -> arr c d -> arr (a, c) (b, d)
(|***|) = parEval2 ()
\end{code}
% \caption{Definition of |parstar|, the parallel version of |***|.}
% \label{fig:par***}
% \end{figure}
% With this we can analogously to the serial |&&&|
We define the parallel |&&&| % (Fig.~\ref{fig:par&&&}) 
in a similar manner to its sequential pendant |&&&| (Fig.~\ref{fig:syntacticSugarArrows}):
% \begin{figure}[h]
\begin{code}
(|&&&|) :: (ArrowChoice arr, ArrowParallel arr (Either a a) (Either b c) ()) =>
	arr a b -> arr a c -> arr a (b, c)
(|&&&|) f g = (arr $ \a -> (a, a)) >>> f |***| g
\end{code} % $ %% formatting
% \caption{Definition of |parand| - the parallel version of |&&&|.}
% \label{fig:par&&&}
% \end{figure}

>>>>

Finally, with the help of |mapArr| (Fig.~\ref{fig:mapArr}), we can define |zipWithArr|  (Fig.~\ref{fig:zipWithArr}) that lifts any Arrow |arr (a, b) c| to an Arrow |arr ([a], [b]) [c]|.
\begin{figure}[h]
\begin{code}
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr (\(as, bs) -> zipWith (,) as bs)) >>> mapArr f
\end{code}
\caption{|zipWith| over Arrows.} %$
\label{fig:zipWithArr}
\end{figure}
 %$ %% formatting

@@

\section{Syntactic Sugar} \label{syntacticSugar}
Finally, we also give the definitions for some syntactic sugar for PArrows, namely |parstar| and |parand|.
For basic arrows, we have the |***| combinator (Fig.~\ref{fig:syntacticSugarArrows}) which allows us to combine two arrows |arr a b| and |arr c d| into an arrow |arr (a, c) (b, d)| which does both computations at once. This can easily be translated into a parallel version |parstar| with the use of |parEval2|, but for this we require a backend which has an implementation that does not require any configuration (hence the |()| as the |conf| parameter% in Fig.~\ref{fig:par***}
):
% \begin{figure}[h]
\begin{code}
(|***|) :: (ArrowChoice arr, ArrowParallel arr (Either a c) (Either b d) ())) =>
	arr a b -> arr c d -> arr (a, c) (b, d)
(|***|) = parEval2 ()
\end{code}
% \caption{Definition of |parstar|, the parallel version of |***|.}
% \label{fig:par***}
% \end{figure}
% With this we can analogously to the serial |&&&|
We define the parallel |parand| % (Fig.~\ref{fig:par&&&}) 
in a similar manner to its sequential pendant |&&&| (Fig.~\ref{fig:syntacticSugarArrows}):
% \begin{figure}[h]
\begin{code}
(|&&&|) :: (ArrowChoice arr, ArrowParallel arr (Either a a) (Either b c) ()) =>
	arr a b -> arr a c -> arr a (b, c)
(|&&&|) f g = (arr $ \a -> (a, a)) >>> f |***| g
\end{code} % $ %% formatting
% \caption{Definition of |parand| - the parallel version of |&&&|.}
% \label{fig:par&&&}
% \end{figure}
